---
title: "Plotting your first map in R"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Connor French"
date: "4/12/2019"
output: html_document
---
## Overview
This tutorial is aimed at folks used to performing GIS tasks on GUI platforms like ArcGIS or QGIS who want to explore R's potential as a powerful, more reproducible, system for spatial visualization and analysis. It assumes no previous command-line experience. The tutorial is organized in the following sections:

**Installing and loading packages**

**Reading in data**

**Plotting the map**

If you haven't already, click [here](https://github.com/connor-french/maps-in-r-tutorial/archive/master.zip) to download the code and data for the tutorial. Just unzip the file, open the .Rmd file if you are using RStudio, or the .R file if you are running things from the command line or in base R. 

## Installing and loading packages
If you need to use a package that isn't part of base R, you first need to install it. The install.packages() function is typically the safest and easiest way to install a package available in the CRAN repository. I put a hashmark in front of the function so it doesn't query the repository every time you want to run this code. Remove the hashmark to install, then you can either remove the line of code, or add a hashmark back in front.

After the package is installed, it's not quite ready to be used. To load a package into your session, use the library() function. 

The main package we're using in this tutorial is the raster package, a set of diverse tools for creating, reading, manipulating, and
writing raster or vector data. This is a wonderful package that can do most, if not all, of the raster manipulation tasks you regularly perform on a platform like ArcGIS or QGIS.
We're also using the rnaturalearth package to load in the shapefile we'll need for our base map.
```{r}
#Install the raster package. You only need to do this once.
#install.packages(c("raster", "rnaturalearth"), dependencies = TRUE)


#Load the raster and rnaturalearth packages. The order you load packages matters, because some packages use similar function names and the last loaded package supercedes priority for function name.
library(raster)
library(rnaturalearth)
```

## Reading in data

Before you can visualize your data, you need to read it in. For this tutorial, we will use two types of data- a shapefile for our base map and latitude/longitude coordinates for our points of interest.

```{r}
#reading in a map of brazil
sa_map <- ne_countries(continent = "south america")

#reading in csv file of the coordinates of the frog localities
frog_locs <- read.csv("data/frog_localities.csv")

```


## Plotting the map

```{r}
plot(sa_map)
points(frog_locs, col = "darkgreen")
```




